---
title: "HGSOC Survival"
author: "Ariel Hippen"
date: '2023-02-18'
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In first_*_analysis.Rmd, we did a preliminary survival analysis of how fibroblast content affects survival, but with the simplest model possible. Here, we will redo that analysis but considering important covariates like age and tumor stage. 

```{r packages}
suppressPackageStartupMessages({
    library(data.table)
    library(SingleCellExperiment)
    library(dplyr)
    library(yaml)
    library(stringr)
    library(ggplot2)
    library(survival)
    library(ggfortify)
})

params <- read_yaml("../../config.yml")
data_path <- params$data_path
local_data_path <- params$local_data_path
plot_path <- params$plot_path
```

## Load data

```{r}
aaces_t <- fread(paste(local_data_path, "deconvolution_output",
                     "AACES_default_bayesprism_results.tsv", sep = "/"))
tcga_t <- fread(paste(local_data_path, "deconvolution_output",
                     "TCGA_default_bayesprism_results.tsv", sep = "/"))
tothill_t <- fread(paste(local_data_path, "deconvolution_output",
                     "tothill_default_bayesprism_results.tsv", sep = "/"))
```

```{r}
# Switch so cell types are columns and samples are rows for easier analysis
cell_types <- aaces_t$cell_type

aaces_t$cell_type <- NULL
aaces <- t(as.matrix(aaces_t))
colnames(aaces) <- cell_types
aaces <- as.data.frame(aaces)
aaces <- cbind(rownames(aaces), aaces)
setnames(aaces, "rownames(aaces)", "ID")
aaces$dataset <- "AA"

tcga_t$cell_type <- NULL
tcga <- t(as.matrix(tcga_t))
colnames(tcga) <- cell_types
tcga <- as.data.frame(tcga)
tcga <- cbind(rownames(tcga), tcga)
setnames(tcga, "rownames(tcga)", "ID")
tcga$dataset <- "TCGA"

tothill_t$cell_type <- NULL
tothill <- t(as.matrix(tothill_t))
colnames(tothill) <- cell_types
tothill <- as.data.frame(tothill)
tothill <- cbind(rownames(tothill), tothill)
setnames(tothill, "rownames(tothill)", "ID")
tothill$dataset <- "tothill"

rm(aaces_t, tcga_t, tothill_t); gc()
```

```{r}
tcga$ID <- str_extract(tcga$ID, "TCGA-\\w\\w-\\w\\w\\w\\w")
tcga$ID <- gsub("-", "\\.", tcga$ID)
```


```{r}
# Combine cell type info into one object to run combined survival analysis
composition <- rbind(aaces, tcga, tothill)
```

```{r}
covariates <- fread(paste(local_data_path, "cluster_assignments",
                          "AnalSet.csv", sep = "/"))
covariates$V1 <- NULL
setnames(covariates, "sampleid", "ID")

covariates$debulking <- recode(covariates$debulking,
                               "Optimal" = "optimal",
                               "Suboptimal" = "suboptimal")

composition <- inner_join(composition, covariates)
```

## AACES

```{r}
aaces <- subset(composition, composition$Dataset=="AA")
```


```{r}
# Discretize fibroblast estimates
quantiles <- quantile(aaces$Fibroblasts)
q1 <- quantiles[2]
median <- quantiles[3]
q3 <- quantiles[4]
aaces$high_fibro <- ifelse(aaces$Fibroblasts > q3, 1, 0)
```

```{r}
# Basic analysis
model <- coxph(Surv(months, vital) ~ (factor(high_fibro)), aaces)
model
```

```{r}
# Analysis with covariates
model <- coxph(Surv(months, vital) ~ (factor(high_fibro)+ factor(age)+ factor(debulking)+ factor(FewerStage)), aaces)
model
```

## TCGA

```{r}
tcga <- subset(composition, composition$Dataset=="TCGA")
```


```{r}
# Discretize fibroblast estimates
quantiles <- quantile(tcga$Fibroblasts)
q1 <- quantiles[2]
median <- quantiles[3]
q3 <- quantiles[4]
tcga$high_fibro <- ifelse(tcga$Fibroblasts > q3, 1, 0)
```

```{r}
# Basic analysis
model <- coxph(Surv(months, vital) ~ (factor(high_fibro)), tcga)
model
```

```{r}
# Analysis with covariates
model <- coxph(Surv(months, vital) ~ (factor(high_fibro)+ factor(age)+ factor(debulking)+ factor(FewerStage)), tcga)
model
```

## Tothill

```{r}
tothill <- subset(composition, composition$Dataset=="Tothill")
```

```{r}
# Discretize fibroblast estimates
quantiles <- quantile(tothill$Fibroblasts)
q1 <- quantiles[2]
median <- quantiles[3]
q3 <- quantiles[4]
tothill$high_fibro <- ifelse(tothill$Fibroblasts > q3, 1, 0)
```

```{r}
# Basic analysis
model <- coxph(Surv(months, vital) ~ (factor(high_fibro)), tothill)
model
```

```{r}
# Analysis with covariates
model <- coxph(Surv(months, vital) ~ (factor(high_fibro)+ factor(age)+ factor(debulking)+ factor(FewerStage)), tothill)
model
```

## All

```{r}
# Discretize fibroblast estimates
quantiles <- quantile(composition$Fibroblasts)
q1 <- quantiles[2]
median <- quantiles[3]
q3 <- quantiles[4]
composition$high_fibro <- ifelse(composition$Fibroblasts > q3, 1, 0)
```

```{r}
# Basic analysis
model <- coxph(Surv(months, vital) ~ (factor(high_fibro)), composition)
model
```

```{r}
# Analysis with covariates
model <- coxph(Surv(months, vital) ~ (factor(high_fibro)+ factor(Dataset)+ factor(age) + factor(debulking) + factor(FewerStage)), composition)
model
```
```{r}
# Kaplan Meier plot for all datasets
km <- Surv(composition$months, composition$vital)
km_treatment<-survfit(km~high_fibro,data=composition,type='kaplan-meier',conf.type='log')

autoplot(km_treatment)
```

Looks like high fibroblast content is associated with worse survival. It doesn't hit statistical significance most of the time, but the results do seem on the same scale as what Mollie saw of the effect of subtypes on survival (https://github.com/greenelab/hgsc_characterization/blob/master/figure_notebooks/survival_figs/survival_analyses.R). It's obviously not going to matter as much as stage or age, but it's still important to consider.